# ProPer clusterAnalysis (VI): find clusters based on sequences of ProPer metrics

```{r clean_start}
rm(list = ls())

## Load required libraries 
require(data.table) # wide format
require(dplyr) # distance matrices
require(scales) # rescale and combine distmats
require(ggplot2) # line plot

#### LOAD DATA ####
read.csv("data_tables/nano_scores_df.csv") -> df
read.csv("data_tables/mini_scores_df.csv")$intervalDuration -> df$dur

```

## Cluster analysis procedure

```{r preparation, warning=FALSE}
# check counts (for checking)
as.data.frame(table(df$file)) -> counts

# remove cases without measures
df[is.na(df$DeltaF0_rel)==F,] -> df
df[is.na(df$sync_rel)==F,] -> df
df[is.na(df$mass_rel)==F,] -> df
df[is.na(df$dur)==F,] -> df

# remove cases without syll label
df[is.na(df$syll_label)==F,] -> df

# remove cases with less than 6 syllables
0 -> df$rm
for (f in unique(df$file)){
  if (nrow(df[df$file==f,])<6){
    1 -> df$rm[df$file==f]
  }
  if (length(unique(df$syll_label[df$file==f]))<6){
    1 -> df$rm[df$file==f]
  }
}
df[df$rm==0,] -> df

# number syllables
df$syll_label -> df$syll_nr
ifelse(df$syll_label=="can"|df$syll_label=="could"|df$syll_label=="may",1,df$syll_nr) -> df$syll_nr
ifelse(df$syll_label=="I",2,df$syll_nr) -> df$syll_nr
ifelse(df$syll_label=="ask",3,df$syll_nr) -> df$syll_nr
ifelse(df$syll_label=="you a",4,df$syll_nr) -> df$syll_nr
ifelse(df$syll_label=="ques-",5,df$syll_nr) -> df$syll_nr
ifelse(df$syll_label=="tion",6,df$syll_nr) -> df$syll_nr
table(df$syll_label)

# wide format
dcast(setDT(df), file ~ syll_nr, value.var = c("DeltaF0_rel","sync_rel")) -> df

# distance matrices
dist(df %>%
       select(grep("Delta", colnames(df)))
) -> dist_delta

dist(df %>%
       select(grep("sync", colnames(df)))
) -> dist_sync

# rescale and combine distmats
rescale(dist_delta,c(0,1)) -> dist_delta
rescale(dist_sync,c(0,1)) -> dist_sync

```

## MAKE TWO CHOICES:

1. which metric to consider, DeltaF0 and/or Synchrony?
Use T to include and F to exclude (both can be T but not F)

2. how many clusters? 
Set 'k' to numbers between 2 and 12

```{r paramaterSetting, warning=FALSE}
# 1. which (combination of) measure(s)?

delta_f0 = T
synchrony = T

# 2. which number of clusters?

k = 6
```

## Run the analysis

```{r justDoIt, warning=FALSE}
# get the (combined) distance matrix
if (delta_f0 == T & synchrony == T){
  dist_delta + dist_sync -> dist
}

if (delta_f0 == T & synchrony == F){
  dist_delta -> dist
}

if (delta_f0 == F & synchrony == T){
  dist_sync -> dist
}

# clustering
hclust(dist,method = "complete") -> hc
cutree(hc,k) -> df$cluster
```

## produce plots

```{r plot, warning=FALSE}
# line plot
melt(df,id.vars = c("file","cluster"),measure.vars = list(c(2:7),c(8:13)),variable.name = "syll_nr",value.name = c("delta","sync")) -> df_plot


clrs <- c("Delta f0"="red", "Synchrony"="blue")
rescale(x = df_plot$sync, to =  c(min(df_plot$delta), max(df_plot$delta))) -> df_plot$sync_resc
a.diff <- max(df_plot$delta) - min(df_plot$delta)
b.diff <- max(df_plot$sync) - min(df_plot$sync)
a.min <- min(df_plot$delta)
b.min <- min(df_plot$sync)

rm(panel_text)
rm(wrap_labs)
panel_text <-
  data.frame(label = paste0("n=", as.character(as.data.frame(
    table(df$cluster)
  )[, 2])),
  cluster = 1:k)
wrap_labs <- (paste0(panel_text$cluster, " (", panel_text$label, ")"))
attributes(wrap_labs)$names <- panel_text$cluster

cues=2

ggplot(df_plot, aes(x = syll_nr))  +
  scale_colour_manual(name="Measure:", values=clrs) +
  scale_y_continuous(
    name = names(clrs[1]),
    sec.axis = sec_axis(~((. -a.min) * b.diff / a.diff) + b.min, name=names(clrs[2]))
  ) +
  theme(legend.position="bottom")  +
  facet_wrap(~ cluster,nrow=2,labeller = as_labeller(wrap_labs)) +
  ylab("") +
  scale_x_discrete(name ="Syllable", labels=c("Can","I","ask","you a","ques-","tion"))+
    theme(plot.title = element_blank(), plot.background = element_rect(fill = "white"), panel.grid.major.x = element_line(), panel.grid.major.y = element_blank(), axis.title.x = element_blank(), axis.title.y = element_text(size = 12), axis.text.x = element_text(size = 9), axis.text.y = element_text(size = 9), strip.text = element_text(size = 12)) -> p
# theme_classic(base_size = 20)-> p

if (synchrony == T){
p +
  stat_summary(
    mapping = aes(y =sync_resc, colour = "Synchrony"),
    fun = mean,
    group = "cluster",
    geom = "line",
    linewidth = 1,
    show.legend = T
  ) -> p}

if (delta_f0 == T){
p +
  stat_summary(
    mapping = aes(y =delta, colour = "Delta f0"),
    fun = mean,
    group = "cluster",
    geom = "line",
    linewidth = 1,
    show.legend = T
  ) -> p}

p
ggsave(p, file=paste0("plots/ClusterLinePlot_",k,"Clusters.pdf"),device=cairo_pdf)

# violin plot
ggplot(df_plot, aes(x = syll_nr)) + 
  scale_colour_manual(name="Measure:", values=clrs) +
  scale_y_continuous(
    name = names(clrs[1]),
    sec.axis = sec_axis(~((. -a.min) * b.diff / a.diff) + b.min, name=names(clrs[2]))
  ) +
  theme(legend.position="bottom")  +
  facet_wrap(~ cluster,nrow=2,labeller = as_labeller(wrap_labs)) +
  ylab("") +
  scale_x_discrete(name ="Syllable", labels=c("Can","I","ask","you a","ques-","tion"))+
    theme(plot.title = element_blank(), plot.background = element_rect(fill = "white"), panel.grid.major.x = element_line(), panel.grid.major.y = element_blank(), axis.title.x = element_blank(), axis.title.y = element_text(size = 12), axis.text.x = element_text(size = 9), axis.text.y = element_text(size = 9), strip.text = element_text(size = 12)) -> vp
  # theme_classic(base_size = 20) -> vp

if (delta_f0 ==T){
vp +   geom_violin(
  mapping = aes(y =delta, colour = "Delta f0"),
  fill = "transparent"
) -> vp}

if(synchrony ==T){
vp +  geom_violin(
    mapping = aes(y =sync_resc, colour = "Synchrony"),
    fill = "transparent"
) -> vp}

vp
ggsave(vp, file=paste0("plots/ClusterViolinPlot_",k,"Clusters.pdf"),device=cairo_pdf)

```

## Select prototypes per cluster

```{r prototypes, warning=FALSE}
# get prototypes per cluster based on the minimum deviation from the cluster mean (iterated over all syllables)
0 -> df$devDlt
0 -> df$devSnc

for (r in 1:nrow(df)){
  mean(unlist(abs(df[r,2:7]-colMeans(df[,2:7][df$cluster==df$cluster[r]])))) -> df$devDlt[r]
  mean(unlist(abs(df[r,8:13]-colMeans(df[,8:13][df$cluster==df$cluster[r]])))) -> df$devSnc[r]
}
rescale(df$devDlt,c(0,1)) -> df$devDlt
rescale(df$devSnc,c(0,1)) -> df$devSnc

for (c in 1:max(df$cluster)){
  print(paste0(c,": ",df$file[df$cluster==c][which.min(df$devDlt[df$cluster==c]+df$devSnc[df$cluster==c])]))
}
```

k = 3
[1] "1: cut-KGO_20210126_230000_ABC7_News_Getting_Answers"
[1] "2: cut-KNTV_20180321_063400_The_Tonight_Show_Starring_Jimmy_Fallon"
[1] "3: cut-KGO_20210610_063500_Jimmy_Kimmel_Live"

k = 6
[1] "1: cut-KGO_20210126_230000_ABC7_News_Getting_Answers"
[1] "2: cut-KNTV_20180321_063400_The_Tonight_Show_Starring_Jimmy_Fallon"
[1] "3: cut-BETW_20170322_070700_The_Daily_Show"
[1] "4: cut-FOXNEWSW_20180220_170000_Outnumbered"
[1] "5: cut-CNBC_20120202_190000_Street_Signs"
[1] "6: cut-CSPAN_20170221_140000_Washington_Journal_News_Headlines_and_Viewer_Calls"

k = 8
[1] "1: cut-KGO_20210126_230000_ABC7_News_Getting_Answers"
[1] "2: cut-KGO_20200318_230000_ABC7_News_400PM"
[1] "3: cut-ALJAZ_20210315_113000_Inside_Story"
[1] "4: cut-CSPAN3_20181015_230600_Discussion_on_the_Opioid_Epidemic_at_Women_for_Trump_Event"
[1] "5: cut-FOXNEWSW_20180220_170000_Outnumbered"
[1] "6: cut-CNBC_20120202_190000_Street_Signs"
[1] "7: cut-CSPAN_20170221_140000_Washington_Journal_News_Headlines_and_Viewer_Calls"
[1] "8: cut-CSPAN3_20210213_005900_American_Artifacts_Jim_Crow_Museum_of_Racist_Memorabilia"

# Write clusterAnalysis table

```{r write_clusterAnalysis_df}
## Write the scores data file

clusterAnalysis_df <- droplevels(subset(df, select = c(file, cluster)))

write.csv(clusterAnalysis_df, "data_tables/clusterAnalysis_df.csv", row.names=FALSE)

```

## Evaluate the optimal number of clusters

```{r evaluation, warning=FALSE}
#### eval w/b ####

# evaluate clustering quality for rounds from 2 to 15 clusters using variance within and between clusters
# within cluster variance expected to be lower with more clusters (better compactness)
# between cluster variance expected to be higher with more clusters (better separation)
# optimum lies where w-var and b-var are the furthest apart for the lowest number of clusters

c() -> df.var
wDlt <- c()
wSnc <- c()

bDlt <- c()
bSnc <- c()

for (r in 2:15){
  cutree(hc,k = r) -> df$cluster
  wcDlt <- c()
  wcSnc <- c()
  for (c in 1:r){
    c() -> colsd_wDlt
      for (m in 2:7) {
        append(colsd_wDlt,sd(unlist(df[df$cluster == c,m,with=FALSE]))) -> colsd_wDlt
      }
      append(wcDlt, mean(colsd_wDlt,na.rm = T)) -> wcDlt
    c() -> colsd_wSnc
      for (m in 8:13) {
        append(colsd_wSnc,sd(unlist(df[df$cluster == c,m,with=FALSE]))) -> colsd_wSnc
      }
    append(wcSnc, mean(colsd_wSnc,na.rm = T)) -> wcSnc
  }
  append(wDlt,mean(c(wcDlt,wcDlt),na.rm = T)) -> wDlt
  append(wSnc,mean(c(wcSnc,wcSnc),na.rm = T)) -> wSnc
  append(bDlt,abs(max(wcDlt,na.rm = T))-abs(min(wcDlt,na.rm = T))) -> bDlt
  append(bSnc,abs(max(wcSnc,na.rm = T))-abs(min(wcSnc,na.rm = T))) -> bSnc
}

as.data.frame(cbind(2:15,wDlt,wSnc,bDlt,bSnc)) -> df.var

# rescale
rescale(df.var$wDlt,c(0,1)) -> df.var$wDlt
rescale(df.var$wSnc,c(0,1)) -> df.var$wSnc
rescale(df.var$bDlt,c(0,1)) -> df.var$bDlt
rescale(df.var$bSnc,c(0,1)) -> df.var$bSnc

eval <- ggplot(df.var, aes(x=V1)) +
  geom_line(aes(y=wDlt),colour="red") +
  geom_line(aes(y=wSnc),colour="blue") +
  geom_line(aes(y=bDlt),colour="red", linetype=2) +
  geom_line(aes(y=bSnc),colour="blue", linetype=2) +
  annotate("text",13,0.5,label = paste0("--- between var","\n","\U2015 within var")) +
  ylab("scaled variance") +
  xlab("N clusters") +
  # theme_classic(base_size = 20)
  theme(plot.title = element_blank(), axis.title = element_text(size = 10), axis.text.x = element_text(size = 9), axis.text.y = element_text(size = 10), strip.text = element_text(size = 8))

print(eval)
ggsave(eval, file=paste0("plots/EvaluationClusterPlot.pdf"),device=cairo_pdf)
```

